\begin{thebibliography}{99}




% \bibitem{RusselNorvig}
%   Stuart Russel, Peter Norvig,
%   \emph{Artificial intelligence, },
%   .


% \bibitem{russel}
%   Stuart Russell, Peter Norvig,
%   \emph{Artificial Intelligence: A Modern Approach, 3rd ed.},
%   Prentice Hall, 2010.

% \bibitem{UrviPatel}
%   Urvi Patel,
%   \emph{Computerized respiratory sound analysis: an exploration of methods},
%   CWRU School of Engineering, Department of Physics, 2011.

% \bibitem{LauraMason}
%   Laura Mason,
%   \emph{Signal processing methods for non invasive respiration monitoring},
%   University of Oxford, 2002,


% \bibitem{Leontios}
%   Leontios Hadjileontiadis,
%   \emph{Lung sounds: an advanced signal processing perspective},
%   Systhesis lectures on biomedical engineering, 2009.

% \bibitem{SCE}
%   Sun X, Cheetham BM, Earis JE,
%   \emph{Real time analysis of lung sounds},
%   Department of Electrical Engineering \& Electronics, Liverpool University, UK.

% ____________________________________________________________________________________________________________________________________ %
% ____________________________________________________________________________________________________________________________________ %
% ____________________________________________________________________________________________________________________________________ %

\bibitem{CACMVS}
  G. Charbonneau, E. Ademovic, B.M.G. Cheetham, L.P. Malmberg, J. Vanderschoot, A.R.A. Sovij\"{a}rvi.
  \emph{Basic techniques for respiratory sound analysis}.

\bibitem{GPA}
  N. Gavriely, Y. Palti, G. Alroy.
  \emph{Spectral characteristics of normal breath sounds}.
  Journal of applied physiology  1981 Feb;50(2):307-14.

\bibitem{SKCAKM}
  S.K. Chowdhury, A.K.Majumder.
  \emph{Digital spectrum analysis of respiratory sounds}.
  IEEE Transactions on Biomedical Engineering. 
  Date of Publication: Nov. 1981. 
  Volume: BME-28, Issue: 11 
  Page(s):784-788 


\bibitem{CORS}
  Simon M\"{u}ller.
  \emph{Classification of Respiratory Sounds}.
  Institut f\"{u}r Stochastik und Anwendungen, Universitaet Stuttgart.
  \url{www.isa.uni-stuttgart.de/LstStoch/Mueller/rbk_17_12_08.pdf}.
  data di accesso 28/01/2013.
  
\bibitem{ADT}
  \url{http://atlantemedicina.wordpress.com/2008/11/04/auscultazione-del-torace/}\\
  data di accesso 28/01/2013.

\bibitem{KoronaKokar}
  Zbigniew Korona, Mieczyslaw M. Kokar.
  \emph{Lung sound recognition using model theory based feature selection and fusion}.
  Applied Signal Processing (October 1998), 5(3), pg. 152-169. 
%   Department of electrical and computer engineering,  Northeastern university 360 Huntington Avenu.
  


\bibitem{kandaswamy}
  C. Kandaswamy.
  \emph{Neural classification of lung sounds using wavelet coefficients}.
  Computer in biology and medicine,  Volume 34, Issue 6, September 2004, Pages 523 - 537.

\bibitem{DOIAERSINTSOTHT}
  Kompis, M. 
  \emph{Distribution of inspiratory and expiratory respiratory sound intensity on the surface of the human thorax}.
  Engineering in Medicine and Biology Society, 1997. Proceedings of the 19th Annual International Conference of the IEEE.

% ____________________________________________________________________________________________________________________________________ %
% ____________________________________________________________________________________________________________________________________ %
% ____________________________________________________________________________________________________________________________________ %


\bibitem{TLSA}
  Rosqvist T, Paajanen E, Kallio K, Rajala HM, Katila T, Piiril\"{a} P, Malmberg P, Sovij\"{a}rvi A.
  \emph{Toolkit for lung sound analysis}.
  Department of Technical Physics, Helsinki University of Technology, Finland.

\bibitem{RSCUCAGMM}
  M. Bahoura, C. Pelletier.
  \emph{Respiratory sounds classification using cepstral analysis and gaussian mixture models}.
  D\'epartement de Math\'ematiques, d'Informatique et de G\'enie (DMIG).
  Universit\`e du Qu\'ebec \`e Rimouski, Rimouski, Qc, Canada, G5L 3A1

\bibitem{ADFMFPLS}
  P.A. Mastorocostas, D.N. Varsamis, C.A. Mastorocostas, C.S. Hilas.
  \emph{A Dynamic Fuzzy Model for Processing Lung Sounds}.
  Department of Informatics and Communications Technological Educational Institute of Serres 62124, Serres, Greece.

\bibitem{ANAAFIORS}
  Feng Jin, Student Member, IEEE and Farook Sattar, Member, IEEE.
  \emph{A new automaed approach for identification of respiratory sounds}.
  School of Electrical and Electronic Engineering, Nanyang Technological University, Nanyang Avenue, Singapore 639798.

\bibitem{PKW}
  Hans Pasterkamp, Steve S. Kraman, George R. Wodicka.
  \emph{Respiratory sounds: advances beyond the stethoscope}.

\bibitem{SoundRepositories}
  Sound repositories:
  \begin{itemize}
    \item 
      \url{http://solutions.3mitalia.it/wps/portal/3M/it_IT/Littmann/stethoscope/education/heart-lung-sounds/}    
    \item
      \url{http://tracheostomy.com/resources/videos/index.htm}
    \item
      \url{http://faemse.org/downloads.html}
    \item
      \url{http://www.meddean.luc.edu/lumen/MedEd/medicine/pulmonar/pd/auditory.htm}
  \end{itemize}

  

\bibitem{AntoniettaBisulli}
  Antonietta Bisulli.
  \emph{Sindrome delle apnee ostruttive nel sonno(osas): effetti cognitivi del trattamento con pressione continua positiva(CPAP)}.
  Dottorato di ricerca. Universit\`a di Bologna.
  
\bibitem{NHLBI}
  NHLBI: Health Information for the Public. 
  \emph{Sleep Apnea: What Is Sleep Apnea?}.
  U.S. Department of Health and Human Services. 

\bibitem{OSARFSD}
  H. Klar Yaggi, M.D., M.P.H., John Concato, M.D., M.P.H., Walter N. Kernan,
  M.D., Judith H. Lichtman, Ph.D., M.P.H., Lawrence M. Brass, M.D., and Vahid Mohsenin, M.D.
  \emph{Obstructive Sleep Apnea as a Risk Factor for Stroke and Death}.
  New England Journal of Medicine 2005; 353:2034-2041N.

\bibitem{DNPSDOSA}
  Apoor S. Gami, M.D., Daniel E. Howard, B.S., Eric J. Olson, M.D., and Virend K. Somers, M.D., Ph.D.
  \emph{Day Night Pattern of Sudden Death in Obstructive Sleep Apnea}.
  New England Journal of Medicine 2005; 352:1206-1214.

\bibitem{PCHDIIRE}
  Christopher Wren, Sam Richmond, Liam Donaldson.
  \emph{Presentation of congenital heart disease in infancy: implications for routine examination}.
  Archive of Disease in Childhood; Fetal and Neonatal Ed. 1999 January; 80(1): F49–F53.

% \bibitem{Forgacs}
%   \emph{Functional Basis of Pulmonary Sounds}
%   Paul Porgacs,
%   journal.publications.chestnet.org

% ____________________________________________________________________________________________________________________________________ %
\bibitem{Pekonen}
  Jussi Pekonen.
  \emph{Onset detection methods for musical sounds}.
  Helsinki university of technology.

\bibitem{Bello}
  Juan Pablo Bello, Laurent Daudet, Samer Abdallah, Chris Duxbury, Mike Davies, and Mark B. Sandler.
  \emph{A Tutorial on Onset Detection in Music Signals}.
% ____________________________________________________________________________________________________________________________________ %

\bibitem{Kauppinen}
  lsmo Kauppinen.
  \emph{Methods for detecting impulsive noise in speech and audio signals}.
  University of Turku, Department of Physics, FIN-ZOOl4 Turku, Finland.


\bibitem{anatomiaRespiratorio}
   \url{ http://pacs.unica.it/biblio/fisiopatologia/fisiopatologia1.pdf}\\
   data di accesso: 28/01/2013.


\bibitem{NSCMBKA}
  Roberto Marani, Gennaro Gelao, and Anna Gina Perri.
  \emph{A New System for Continuous Monitoring of Breathing and Kinetic Activity}.
  Electrical and Electronic Department, Polytechnic of Bari, Via E. Orabona, Bari 70125, Italy.

\bibitem{Invasivita}
  \url{http://it.wikipedia.org/wiki/Invasivit\`a}\\
  data di accesso: 28/01/2013.


\bibitem{ASDBOS}
  Arzt M, Young T, Finn L, Skatrud JB, Bradley TD.
  \emph{Association of sleep disordered breathing and the occurrence of stroke}.
  Toronto General Hospital, University Health Network, 9N 943, 200 Elizabeth Street, Toronto, ON M5G 2C4, Canada.

\bibitem{ASPODUOCSS}
  Yildirim I, Ansari R, Moussavi Z.
  \emph{Automated respiratory phase and onset detection using only chest sound signal}.
  University of Illinois at Chicago, IL 60607, USA.


\bibitem{BARFUTS}
  Saiful Huq, Azadeh Yadollahi, Zahra Moussavi.
  \emph{Breath Analysis of Respiratory Flow using Tracheal Sounds}.
  Department of Electrical and Computer Engineering, University of Manitoba, Winnipeg, MB, Canada.


%_______________________________________________________________________________________________________________________________________$

\bibitem{PneumotacofragoTreCani}
  \url{http://www.treccani.it/vocabolario/pneumotacografo/}\\
  data di acesso: 29/01/2013.
\bibitem{WikiSpirPneu}
  \url{http://it.wikipedia.org/wiki/Spirometro#Spirometro_con_pneumotacografo}\\
  data di acesso: 29/01/2013.

\bibitem{polisonnografo}
  \url{http://www.nlm.nih.gov/medlineplus/ency/article/003932.htm}\\
  data di accesso: 04/02/2013.
%_______________________________________________________________________________________________________________________________________$

\bibitem{TFC}
  Tiago H. Falk and Wai Yip Chan.
  \emph{Modulation filtering for heart and lung sound separation from breath sound recordings}.
  Department of Electrical and Computer Engineering, Queen's University, Canada.

\bibitem{ARSAPD}
  \emph{Acoustical respiratory signal analysis and phase detection}.
  S. Le Cam, Ch. Collet, F. Salzenstein.
  Universit\'e Strasbourg 1.

\bibitem{SCRSA}
  A.R.A. Sovij\"{a}rvi, J. Vanderschoot, J.E. Earis.
  \emph{Standardization of computerized respiratory sound analysis}.





%__________________________________________________________________________________$

\bibitem{RSDUVFD}
  Yee Leng Yap, Zahra Moussavi,
  \emph{Respiratory onset detection using variance fractal dimension},
  Department of Electrical Engineering, University of Manitoba.

\bibitem{FLSCTWFDA}
  January Gnitecki, Zahra Moussavi.
  \emph{The fractality of lung sounds: A comparison of three waveform fractal dimension algorithms}.
  Faculty of Engineering, Department of Electrical and Computer Engineering, University of Manitoba, Winnipeg, MB Canada R3T 5V6.

%__________________________________________________________________________________$



\bibitem{DECE}
  Joo S. Chuah , Zahra K. Moussavi.
  \emph{Automated respiratory phase detection by acoustical means}.
  Department of Electrical and Computer Engineering, University of Manitoba, Winnipeg, Manitoba, R3T 2N2, Canada.
  

\bibitem{CARPDWAM}
  Z. Moussavi, M. Leopando, H. Pasterkamp, G. Rempel.
  \emph{Computerized acoustical respiratory phase detection without airflow measurament}.
  Medical and Biological Engineering and Computing 2000, Volume 38, Issue 2, pp 198-203.


\bibitem{PDUCABS}
  H. A. Mansy, T. J. Royston, R. A. Balk, R.H. Sandier.
  \emph{Pneumothorax detection using computerised analysis of breath sounds}.
  Medical and Biological Engineering and Computing 2002, Volume 40, Issue 5, pp 526-532.


\bibitem{SPMNIRM}
  Laura Mason.
  \emph{Signal Processing Methods for Non-Invasive Respiration Monitoring}.
  Trinity College Michaelmas 2002.


\bibitem{ASTFARA}
  Gina Ann Yi.
  \emph{A Software Toolkit for Acoustic Respiratory Analysis}.
  Massachussetts institute of technology.

\bibitem{TFTAAOMA}
  Stephen Webley Hainsworth.
  \emph{Techniques for the Automated Analysis of Musical Audio}.
  Magdalene College December 2003.

\bibitem{BP}
  M J Tobin; T S Chadha; G Jenouri; S J Birch; H B Gazeroglu; M A Sackner.
  \emph{Breathing patterns. 1. Normal subjects.}.
  Chest Journal. 1983;84(2):202-205.

\bibitem{ABP}
  \url{http://www.meddean.luc.edu/lumen/MedEd/medicine/pulmonar/physio/pf11.htm}\\
  data di accesso: 01/02/2013.

\bibitem{ASAHAIS}
  Susan Redline1, Gayane Yenokyan2, Daniel J. Gottlieb3,4, Eyal Shahar5, George T. O'Connor3, Helaine E. Resnick6,7, Marie Diener-West2, Mark H. Sanders8, Philip A. Wolf3, Estella M. Geraghty9, Tauqeer Ali9, Michael Lebowitz11, and Naresh M. Punjabi.
  \emph{Obstructive Sleep Apnea Hypopnea and Incident Stroke}.

\bibitem{SSAAROISITE}
  Roberto Munoz, Joaqu\'in Duran-Cantolla, Eduardo Mart\'inez-Vila, Jaime Gallego, Ram\'on Rubio, Felipe Aizpuru and Germ\'an De La Torre.
  \emph{Severe Sleep Apnea and Risk of Ischemic Stroke in the Elderly}.


% \bibitem{PCDP}
%   Ben-Ari, M.,
%   \emph{Principles of Concurrent and Distributed Programming}, 
%   Prentice Hall, 1990. 
% 
%   \bibitem{SAMHRTE} 
%     C. Liu, J. Layland, 
%     \emph{Scheduling Algorithms for Multiprogramming in a Hard Real-time Environment}, 
%    Journal of the ACM, 20(1):46--61, Jan. 1973. http://citeseer.ist.psu.edu/liu73scheduling.html

\bibitem{RealTime}
  \url{www.cis.upenn.edu/~lee/06cse480/lec-real-time-scheduling.pdf}\\
  data di accesso: 02/02/2013.
\bibitem{WikiRealTime}
  \url{http://en.wikipedia.org/wiki/Real-time}\\
  data di accesso: 02/02/2013.

\bibitem{RTDSPIA}
  S.M. Kuo, B.H. Lee, and W. Tian.
  \emph{Real-Time Digital Signal Processing: Implementations and Applications}.
  Wiley, 2006. ISBN 0-470-01495-4.


\bibitem{appaRespi}
  \url{http://it.wikipedia.org/wiki/File:Respiratory_system_complete_it.svg}\\
  data di accesso: 04/02/2013.

\bibitem{wikiMeccaRespi}
  \url{http://it.wikipedia.org/wiki/Respirazione}\\
  data di accesso: 04/02/2013.

\bibitem{meccanicaRespi}
  Ronald B. George, Richard A. Matthay, Michael A. Matthay, Richard W. Light.
  \emph{Chest Medicine: Essentials of Pulmonary and Critical Care Medicine. Edition 5}.
   Lippincott Williams and Wilkins.

% \bibitem{auscu}
%   Francesca Palma,
%   \emph{Classificazione funzionale di valvole cardiache meccaniche attraverso fonocardiografia ad ultrasuoni}
%   Tesi di laurea magistrale in bioingegneria,
%   Universit\'a degli studi di Padova,  2010/2011.




\bibitem{BBATMBS}
  \v Zilbert Tafa, Radovan Stojanovi\'c.
  \emph{Bluetooth based approach to monitoring biomedical signals}.
  Proceedings of the 5th WSEAS International Conference on Telecommunications and Informatics, Istanbul, Turkey, May 27-29, 2006 (pp415-420).


\bibitem{jstk}
  \url{http://code.google.com/p/jstk/}\\
  data di accesso: 04/02/2013.

\bibitem{clientServer}
  \url{http://en.wikipedia.org/wiki/Client-server_model}\\
  data di accesso: 04/02/2013.

\bibitem{javaSerializable}
  \url{http://docs.oracle.com/javase/1.5.0/docs/guide/serialization/spec/serial-arch.html#6428}\\
  data di accesso: 04/02/2013.

\bibitem{clustering}
  Glenn Fung.
  \emph{A Comprehensive Overview of Basic Clustering Algorithms}.
  June 22, 2001.


\bibitem{LSAITTIOTAF}
  S.S. Kraman.
  \emph{Lung Sounds: An Introduction to the Interpretation of the Auscultatory Finding}.
  Northbrook, IL: Amer. College of Chest Physicians, 1993, audio tape.

\bibitem{PatternRecognition}
  Sergios Theodoridis, Konstantinos Koutroumbas.
  \emph{Pattern recognition}.
  Elsevier Academic Press.

\bibitem{envelope}
  C. Richard Johnson, Jr, William A. Sethares, Andrew G. Klein.
  \emph{Software Receiver Design: Build Your Own Digital Communication System in Five Easy Steps}.
  Cambridge University Press. p. 417.

\bibitem{envelopeImg}
  \url{http://en.wikipedia.org/wiki/File:Signal_envelopes.png}\\
  data di accesso: 04/02/2013.

\bibitem{filtri}
  Belle A. Shenoi.
  \emph{Introduction to digital signal processing and filter design}.
  John Wiley and Sons.

\bibitem{intrrr}
  \emph{A Neural Network System for Detection of Obstructive Sleep Apnea Through SpO2 Signal Features}

\bibitem{markovtriple}
  \emph{Cha\v ines de Markov Triplet}.
  Wojciech Pieczynski.

\bibitem{markovtriple2}
  \emph{Multisensor triplet Markov chains and theory of evidence}.
  Wojciech Pieczynski.


\bibitem{fourier}
  Albert Boggess, Francis J. Narcowich.
  \emph{A first course in wavelets with Fourier analysis}.
  
\bibitem{fourier2}
  Paolo Marcellini, Carlo Sbordone.
  \emph{Analisi matematica uno}.

\bibitem{fourier3}
  \url{http://it.wikipedia.org/wiki/Fourier-Transformation}.


\bibitem{wavelet1}
  Rami Cohen.
  \emph{Signal Denoising Using Wavelets}.
  Department of Electrical Engineering Technion, Israel Institute of Technology.


\bibitem{wavelet2}
  R.J.E. Merry.
  \emph{Wavelet Theory and Applications. A literature study}.
   
\bibitem{DomTF}
  Analisi dei segnali campionati.
  \url{www.diee.unica.it/misure/Dispense/Misure_Elettroniche_dm270/Analisi_di_segnali_campionati.pdf}.
  data di accesso: 04/02/2013.

\bibitem{EAP}
  Carlo Drioli, Nicola Orio,
  \emph{Elementi di Acustica e Psicoacustica}


\bibitem{elabA}
  Corso di elaborazione audio.
  \url{http://www.mediasystemnet.it/videocorsi.html}.
  data di accesso: 04/02/2013.


\bibitem{IONDASON}
  \url{http://editoria.wiki-site.com/index.php/DIGITALIZZAZIONE}.
  data di accesso: 04/02/2013.

\bibitem{IDOMTF}
  \url{http://fisicaondemusica.unimore.it/Teorema_di_Fourier.html}.
  data di accesso: 04/02/2013.


\bibitem{SELET}
  \url {http://www.suonoelettronico.com}.
  data di accesso: 04/02/2013.

\bibitem{fusello}
  \emph{Gli stetoscopi elettronici}.  
  Giorgio Carlo Monti, Massimo Fusello.
  \url{http://www.simg.it/Documenti/Rivista/2002/08-10_2002/12.pdf}.
  data di accesso: 04/02/2013.

\bibitem{Swing}
  \url{http://docs.oracle.com/javase/6/docs/technotes/guides/swing/}.
  data di accesso: 04/02/2013.

\bibitem{jmathplot}
  \url{http://code.google.com/p/jmathplot/}.
  data di accesso: 04/02/2013.

\bibitem{audacity}
  \url{http://audacity.sourceforge.net/}
  data di accesso: 04/02/2013.


\bibitem{separazione1}
  \emph{Blind source extraction of heart sound signals from lung soud recordings exploiting periodicity of the heart sound}.
  T. Tsalaile, S. M. Naqvi, K. Nazarpour, S. Sanei and J. A. Chambers.

\bibitem{separazione2}
  \emph{Heart sounds separation from lung sounds using independent component analysis}.
  M. T. Pourazad, Z. Moussavi, F. Farahmand, R. K. Ward.
  Proceedings of the 2005 IEEE Engineering in Medicine and Biology 27th Annual Conference Shanghai, China, September 1-4, 2005.


\bibitem{separazione3}
  \emph{Modulation filtering for heart and lung sound separation from breath sound recordings}.
  Tiago H. Falk and Wai-Yip Chan.
  International Journal of Computer and Electrical Engineering, Vol. 2, No. 3, June, 2010 1793-8163.


\bibitem{separazione4}
  \emph{Separating Heart Sound from Lung Sound Using LabVIEW}
  T E Ayoob Khan, P Vijayakumar.


\bibitem{separazione5}
  \emph{Separating Heart Sounds from Lung Sounds}
  January Gnitecki, Zahara M. K. Moussavi.



\bibitem{separazione6}
  \emph{Separation of heart sound signal from lung sound signal by adaptive line enhancement}
  Thato Tsalaile, Saeid Sanei.
  15th European Signal Processing Conference (EUSIPCO 2007), Poznan, Poland, September 3-7, 2007, copyright by EURASIP.






\end{thebibliography}

